Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra

### 2015-06-05 ms.tex
Update response to referee, minor changes to draft
From 7e21d9e:
============

In response_to_ApJL_referee_report_1.txt,

(*) Update overall and specific responses to the referee report

In ms.tex,

(*) Fix minor typo in Abstract
(*) Minor change in wording in Introduction
(*) Minor change in hyphenation in Section 4
(*) Minor change in wording in Section 4

We confirm our recent prediction of the ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures from the simulation which predicted this feature were validated against Murchison Widefield Array (MWA) data but this key {\it pitchfork} signature was close to the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can by mitigated by careful antenna designs that suppress sensitivity near the horizon. Simple models for antenna apertures proposed for future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array indicate they should suppress foreground leakage from the {\it pitchfork} by $\sim 40$~dB relative to the MWA, and significantly increase the likelihood of cosmological signal detection in these critical Fourier modes in the three-dimensional power spectrum.

@@

Our recent study \citep[][hereafter referred to as Paper~I]{thy15} used instrument and foreground models, for the first time with full sky coverage, in order to simulate actual EoR experiments more accurately than previous studies. Surprisingly, we found that foreground emission outside the primary beam field of view caused the most significant contamination of Fourier modes considered most sensitive for detecting the cosmological H~{\sc i} signal in delay spectrum based analyses. This contamination is the result of the interplay between foreground emission, particularly diffuse Galactic emission, and the wide-field properties typical of EoR instruments. Our simulations predicted that delay spectra from the MWA and other experiments should exhibit a characteristic ``pitchfork'' appearance with local maxima near the horizon delay limits, in addition to at the primary lobe region.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{3_baseline_comparison_CLEAN_fhd_avg_visibilities_amplitudes_185.0_MHz_30.7_MHz.eps}
\includegraphics[width=\linewidth]{f2.eps}
\caption{Delay power spectra on three antenna spacings oriented northward, obtained by coherent averaging of 12 snapshots aligned in LST. The averaged data and models are shown in black and red respectively. The antenna spacings are 83.5~m ({\it top}), 104.2~m ({\it middle}), and 171~m ({\it bottom}). The horizontal dotted line is the {\it rms} of thermal fluctuations. The vertical dashed lines denote the horizon delay limits. The vertical dotted-dashed lines at $\tau = \pm 0.78\,\mu$s correspond to grating responses of periodic flagging of bandpass at intervals of 1.28~MHz. The peaks close to the horizon delay limits are distinctly visible at $\sim 10$-1000~$\sigma$ levels. Differences between model and data are primarily attributed to uncertainties in the foreground model and the MWA tile power pattern. \label{fig:3-baseline-comparison-delay-spectra}}
\end{figure}

We focus on the prominent peaks in data near the horizon limits. Typically, the power near the negative horizon limit is seen with a signal-noise ratio $\sim$~10-100, while that around the positive horizon limit is $\sim$~100-1000.

@@

We note that reducing uncertainties will only change the relative strength of the {\it pitchfork} signature in our model. However, the effects giving rise to this signature are generic to all wide-field measurements. Thus, the extremely high significance detection of foreground emission near the horizon limits is a robust confirmation of the predicted effects of wide-field measurements.

>>>>

We confirm our recent prediction of the ``pitchfork'' foreground signature in power spectra of high-redshift 21~cm measurements, wherein the interferometer is sensitive to large-scale structure on all baselines. This is due to the inherent response of a wide-field instrument and is characterized by enhanced power from foreground emission in Fourier modes adjacent to those considered to be most sensitive to the cosmological H~{\sc i} signal. In our recent paper, many signatures from the simulation which predicted this feature were validated against Murchison Widefield Array (MWA) data but this key {\it pitchfork} signature was close to the noise level. In this paper, we improve the data sensitivity through coherent averaging of 12 independent snapshots with identical instrument settings, and provide the first confirmation of the prediction with a signal-noise ratio $>10$. This wide-field effect can be mitigated by careful antenna designs that suppress sensitivity near the horizon. Simple models for antenna apertures proposed for future instruments such as the Hydrogen Epoch of Reionization Array and the Square Kilometre Array indicate they should suppress foreground leakage from the {\it pitchfork} by $\sim 40$~dB relative to the MWA, and significantly increase the likelihood of cosmological signal detection in these critical Fourier modes in the three-dimensional power spectrum.

@@

Our recent study \citep[][hereafter referred to as Paper~I]{thy15} used instrument and foreground models, for the first time with full sky coverage, in order to simulate actual EoR experiments more accurately than previous studies. Surprisingly, we found that foreground emission outside the primary beam field of view caused the most significant contamination of Fourier modes considered most sensitive for detecting the cosmological H~{\sc i} signal in delay spectrum based analyses. This is the result of the interplay between foreground emission, particularly diffuse Galactic emission, and the wide-field properties typical of EoR instruments. Our simulations predicted that delay spectra from the MWA and other experiments should exhibit a characteristic ``pitchfork'' appearance with local maxima near the horizon delay limits, in addition to at the primary lobe region.

@@

\begin{figure}[htb]
\centering
% \includegraphics[width=\linewidth]{3_baseline_comparison_CLEAN_fhd_avg_visibilities_amplitudes_185.0_MHz_30.7_MHz.eps}
\includegraphics[width=\linewidth]{f2.eps}
\caption{Delay power spectra on three antenna spacings oriented northward, obtained by coherent averaging of 12 snapshots aligned in LST. The averaged data and models are shown in black and red respectively. The antenna spacings are 83.5~m ({\it top}), 104.2~m ({\it middle}), and 171~m ({\it bottom}). The horizontal dotted line is the {\it rms} of thermal fluctuations. The vertical dashed lines denote the horizon delay limits. The vertical dotted-dashed lines at $\tau = \pm 0.78\,\mu$s correspond to grating responses of periodic flagging of bandpass at intervals of 1.28~MHz. The peaks close to the horizon delay limits are distinctly visible at $\sim 10$--1000~$\sigma$ levels. Differences between model and data are primarily attributed to uncertainties in the foreground model and the MWA tile power pattern. \label{fig:3-baseline-comparison-delay-spectra}}
\end{figure}

We focus on the prominent peaks in data near the horizon limits. Typically, the power near the negative horizon limit is seen with a signal-noise ratio $\sim$~10--100, while that around the positive horizon limit is $\sim$~100--1000.

@@

We note that reducing uncertainties will only change the relative strength of the {\it pitchfork} signature in our model. However, the effects giving rise to this signature are generic to all wide-field measurements. Thus, the extremely high-significance detection of foreground emission near the horizon limits is a robust confirmation of the predicted effects of wide-field measurements.
